---
title: "DataGeneration"
author: "Kristen Hayward"
date: '2020-01-24'
output: html_document
---

##PART I

1. Created a public repository on GitHub called Krishayward/BIOL812-ASSIGNMENT2 (https://github.com/Krishayward/BIOL812-ASSIGNMENT2.git)

2. Created a new project linked to my new github directory.

3. Created a new Rmd script called DataGeneration.rmd here.

4. Create an object called Species that is a character vector containing the names of 10 species:
```{r}
Species<- as.character(c("polar bear","ringed seal","bearded seal", "northern lemming","walrus","grizzly bear","american black bear", "koala","arctic fox","homo sapiens"))
Species #check it
```

5. Create a vector called Limb.Width that is a numeric vector containing 10 random numbers sampled from a normal distribution with a mean of 10 and standard deviation of 10:
```{r}
Limb.Width<- rnorm(10, mean = 10, sd = 10)
Limb.Width #check er out
```

6. Change any negative numbers in Limb.Width to positive values:
```{r}
Limb.Width<- abs(Limb.Width)
Limb.Width #make sure it worked
```

7. Create a vector called unitsW that is a boolean vector containing 10 strings that are either mm or cm generate randomly with a 0.5 probabibility of each:
```{r}
unitsW<- sample(c("cm","mm"), 10, replace = TRUE, prob = c(0.5,0.5))
unitsW
```

8. Repeat step 5 & 6 for a new object called Limb.Length, but with a mean and standard deviation of 100:
```{r}
#step 5
Limb.Length<- rnorm(10, mean = 100, sd = 100)
Limb.Length
#step 6
Limb.Length<- abs(Limb.Length)
Limb.Length
```

9. Repeat step 7 for a new object called unitsL:
```{r}
unitsL<- sample(c("cm","mm"), 10, replace = TRUE, prob = c(0.5,0.5))
unitsL
```

10. Combine all vectors into a data.frame object called MyData:
```{r}
MyData<- data.frame(Species, Limb.Width, unitsW, Limb.Length, unitsL)
MyData
```

11. Save MyData as a csv file in the Data folder called measurements.csv:
```{r}
write.csv(MyData, "Data/measurements.csv")
```









