---
title: "DataAnalysis"
author: "Kristen Hayward"
date: '2020-01-24'
output: html_document
---

##PART II

Load packages:
```{r}
library(dplyr)
library(ggplot2)
library(knitr)
```


1. Read measurements.csv into an object in R called MData:
```{r}
MData<- read.csv("Data/measurements.csv", header = T)
MData
```

2. Include functions from the Data Science Tutorial to do each of the following:

a) Change the column names to shorter, 3-letter abbreviations:
```{r}
#use rename to change column titles - New name = old name
MData<- rename(MData, Spp = Species, LbW = Limb.Width,
               UnW = unitsW, LbL = Limb.Length, 
               UnL = unitsL)
```

b) Add a new data column called Volume that calculates limb volume:
```{r}
#first I'm going to make sure all the units are the same - use ifelse function to change all those not in cm to cm 
MData$LbW<- ifelse(MData$UnW == "cm", 1, 0.1)*MData$LbW
MData$LbL<- ifelse(MData$UnL == "cm", 1, 0.1)*MData$LbL
MData$UnW<- "cm"
MData$UnL<- "cm"

#make a new column with volume - limb is cylinder which has the formula V = pi*r2*h
#I used the limb length as the height & the limb width as the diameter of the cylinder
MData<- mutate(MData, Vol = pi*((LbW/2)^2)*LbL)
```

c) Group the species into 2 or more imaginary groups & add a column denoting these groups. I decided to separate the animals into two groups depending on their habitat (terrestrial or marine):
```{r}
#add column using the mutate function from dplyr
MData<- mutate(MData, Hab = c("terrestrial","marine",
              "marine","terrestrial",
              "marine","terrestrial",
              "terrestrial","terrestrial",
              "terrestrial","terrestrial"))
#make new dataset with data grouped by my new habitat variable
SppHab<- group_by(MData,Hab)
SppHab
```

d) Summarize the mean and SD length, width, and volume for each of my groups & present in table form:
```{r}
#use summarize function to generate means & SDs
SumGroup<- summarize(SppHab, mean_L = mean(LbL), SD_L = sd(LbL),
          mean_W = mean(LbW), SD_W = sd(LbW),
          mean_V = mean(Vol), SD_V = sd(Vol))
SumGroup
```
Here is my lovely table:

Species Habitat | Mean limb length (cm) | SD of limb length (cm) | Mean limb width (cm) | SD of limb width (cm) | Mean limb volume (cm^3^) | SD of limb volume (cm^3^)
-----------|--------|--------|--------|--------|--------|--------|
marine | 86.96 | 73.85 | 4.23 | 5.37 | 180.40 | 90.72 
terrestrial | 99.58 | 84.80 | 7.34 | 7.30 | 5730.05 | 7695.12

However, can also make table kable function:
```{r}
#round it 
SumGroup$mean_L<- round(SumGroup$mean_L, digits = 2)
SumGroup$mean_W<- round(SumGroup$mean_W, digits = 2)
SumGroup$mean_V<- round(SumGroup$mean_V, digits = 2)
SumGroup$SD_L<- round(SumGroup$SD_L, digits = 2)
SumGroup$SD_W<- round(SumGroup$SD_W, digits = 2)
SumGroup$SD_V<- round(SumGroup$SD_V, digits = 2)

kable(SumGroup, caption = "Table 1. Mean and standard deviations for limb width, length, and volume for terrestrial and marine species (n = 10).")
```

e) Plot width and length BUT embed the graph in R markdown file rather than creating separate PDF. Here, I just plot limb length against width to visualize a potential correlation:
```{r}
library(ggplot2)
#load the custom theme
cooltoo<- theme(
  legend.position= "none",
  axis.text.x= element_text(size=12, colour="black", face= "bold"),
  axis.text.y= element_text(size=12, colour="black", face= "bold"),
  axis.title.x= element_text(size=14, colour="black", vjust= +0.5, face= "bold"),
  axis.title.y= element_text(size=14, colour="black", vjust=+1, face= "bold"),
  panel.background= element_rect(fill="white", colour=NULL),
  axis.line= element_line(colour="black", size=0.5),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
  )
#create the plot
ggplot(MData, aes(x = LbW, y = LbL))+
  geom_point(size = 4, colour = "sienna4")+ #alter size & colour of data points
  scale_x_continuous(name = "Limb width (cm)")+ #name the axes
  scale_y_continuous(name = "Limb length (cm)")+
  geom_smooth(method = "lm", se = FALSE, colour = "gray15", lwd = 1.5)+ #add a regression line 
  cooltoo #add the custom theme 
```

f) Plot width and length, but only for species that have above-average volume:
```{r}
#select only the species with volume above the mean volume - use filter function to subset 
AboveAVG<- filter(MData, Vol > mean(MData$Vol))

#now plot limb length vs limb width (like above) but only for these species
ggplot(AboveAVG, aes(x = LbW, y = LbL))+
  geom_point(size = 4, colour = "thistle4")+ 
  scale_x_continuous(name = "Limb width (cm)")+ 
  scale_y_continuous(name = "Limb length (cm)")+
  geom_smooth(method = "lm", se = FALSE, colour = "gray15", lwd = 1.5)+ 
  cooltoo 
```
Only three points - not very interesting!

g) Add a table showing only species and volume for only the species with below-average volume:
```{r}
#select only species with below average volume - use filter function to subset
BelowAVG<- filter(MData, Vol < mean(MData$Vol))

#round the digits
BelowAVG$Vol<- round(BelowAVG$Vol, digits = 3)

#make it into a dataframe for table 
BelowAVG<- data.frame("Species" = BelowAVG$Spp,
                      "Volume (cm^3)" = BelowAVG$Vol)

#use kable function to make table or create in markdown text (below):
kable(BelowAVG, caption = "Table 2. Species and their respective limb volumes in cm^3 for species with below average limb volumes from a random sample (n=10).")
```

